Reduced Nucleotide Diversity on a Plant Y Chromosome Following Recent Recombination Suppression

### 2016-08-02 manuscript.tex
finished section on Ne predictions

To consider the joint effects of biased sex ratios and variance in male reproductive success, our neutrally-expected levels of diversity were calculated assuming no variance in female mating (e.g., $C^2_{f}=0$) and we considered a range of values for $\sigma_{m}$: 0 (no variance), 1/2 (Poisson/2) 1 (Poisson distribution of offspring), 2 (Poisson*2), and 3 (Poisson*3). Assuming the level of neutral polymorphism is given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}, Figure 1 shows the predicted X/A and Y/A diversity ratios as a function of the sex ratio and variance in male reproductive success. Note that we have only considered variance in male reproductive success because this is most likely to occur in wind pollinated plant like \textit{R. hastatulus}, where males produce copious amounts of pollen and compete to fertilize the uniovulate females.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=\linewidth]{Figure1.png}
\caption{\textbf{A:} The relation between relative effective population size and sex ratio bias for genes on autosomes (A), X chromosomes (X), and Y chromosomes (Y). The sex ratio is shown as the proportion of females, $N_{f}/(N_{f}+N_{m})=0$, ranging from 0 to 1, where $N_{m}$ and $N_{f}$ are the effective number of breeding males and females, respectively. Calculations assume a constant population size, non-overlapping generations, and a Poisson distribution of offspring number \citep{wright1931evolution,charlesworth2009effective}. The $N_{e}$ for each chromosome type is calculated relative to the predicted $N_{e}$ for genes on autosomes with an equal sex ratio.\textbf{B:} The corresponding expected relation between sex ratio bias and the sex-chromosome-to-autosome diversity ratio, assuming $\theta=4N_{e}\mu$ and equal neutral mutation rates for all genes.
}%
\label{fig:spectrum}
\end{figure}

>>>>

To consider the joint effects of biased sex ratios and variance in male reproductive success, our neutrally-expected levels of diversity were calculated assuming no variance in female mating (e.g., $C^2_{f}=0$) and we considered a range of values for $\sigma_{m}$: 0 (no variance), 1/2 (Poisson/2) 1 (Poisson distribution of offspring), 2 (Poisson*2), and 3 (Poisson*3). Assuming the level of neutral polymorphism is given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}, Figure 1 shows the predicted X/A and Y/A diversity ratios as a function of the sex ratio and variance in male reproductive success. Note that we have only considered variance in male reproductive success because this is most likely to occur in wind pollinated plant like \textit{R. hastatulus}, where males produce copious amounts of pollen and compete to fertilize uniovulate females.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=\linewidth]{Figure1.png}
\caption{The relation between relative effective population size and sex ratio bias for genes on autosomes (\textbf{A}) and sex chromosomes (\textbf{B}; Y chromosome in red). The sex ratio is shown as the proportion of males, $N_{m}/(N_{f}+N_{m})$, where $N_{m}$ and $N_{f}$ are the effective number of breeding males and females, respectively, plotted against $N_{e}/N$, where $N=N_{m}+N_{f}$. Dotted lines show predictions where both males and females produce Poisson-distributed offspring numbers. In that case, the $N_{e}$'s are given by $N_{e{A}}=4N_{m}N_{f}/(N_{m}+N_{f})$ for autosomes, $N_{e{X}}=9N_{m}N_{f}/(4N_{m}+2N_{f})$ for the X, and $N_{e{Y}}=N_{m}/2$ for the Y \citep{wright1931evolution}. The other curves correspond to increasing levels of variance in male reproductive success \citep{nomura2002effective} (see Methods). Assuming $\theta=4N_{e}\mu$ and equal neutral mutation rates among genes, the predicted $N_{e}$'s are used to generate our null expectations for the X/A and Y/A ratios of diversity.  
}%
\label{fig:spectrum}
\end{figure}
